---
title: "cars"
author: "Yadu"
date: "July 23, 2015"
output: html_document
---
First we have to install the package ggplot2:
```{r}
#install.packages("ggplot2")
```

Next, we have to load the following packages (ggplot2 to generate scatterplots, plyr to change the column names, and datasets so that we can obtain the mtcars dataset:
```{r}
library(ggplot2)
library(plyr)
library(datasets)
```

Next, we change the names of the columns as follows:
```{r}
colnames(mtcars) <- c("Mileage (miles/gallon)", 
"Number of Cylinders", 
"Displacement (cubic inches)", 
"Gross Horsepower", 
"Rear Axle Ratio", 
"Weight (lb/1000)",
"Quarter Mile Time (Seconds)", 
"Engine Shape (V or Straight)", 
"Manual or Automatic Transmission",
"Number of Forward Gears", 
"Number of Carburettors")
mtcars
```

Next, we have to substitute some of the quantitative data with qualitative data:
```{r}
mtcars$`Engine Shape (V or Straight)` <- gsub("0", "V", 
mtcars$`Engine Shape (V or Straight)`)
mtcars$`Engine Shape (V or Straight)` <- gsub("1", "Straight", 
mtcars$`Engine Shape (V or Straight)`)
mtcars$`Manual or Automatic Transmission` <- gsub("0", "Automatic", 
mtcars$`Manual or Automatic Transmission`)
mtcars$`Manual or Automatic Transmission` <- gsub("1", "Manual", 
mtcars$`Manual or Automatic Transmission`)
mtcars
```

Next, we obtain data regarding the internal structure of the data:
```{r}
str(mtcars)
summary(mtcars)
```

Next, we calculate the amount of time required to travel a full mile as follows:
```{r}
mtcars$`Full Mile Time (Seconds)` <- (mtcars$`Quarter Mile Time (Seconds)`)*4
```

The result is an entire new column consisting of quantitative data:
```{r}
mtcars$`Full Mile Time (Seconds)`
```

Next, we take the inverse of the quantity previously calculated to obtain the maximum speed of the vehicle:
```{r}
mtcars$`Speed (Miles Per Second)` <- 1/(mtcars$`Full Mile Time (Seconds)`)
```

Here also, we obtain an entire new column consisting of quantitative data:
```{r}
mtcars$`Speed (Miles Per Second)`
```

Next, we calculate the number of revolutions per minute as follows:
```{r}
mtcars$`Revolutions Per Minute` <- (mtcars$`Speed (Miles Per Second)`)*2112*(mtcars$`Rear Axle Ratio`)
```

Like the previous two data sets, this is also a column consisting of quantitative data:
```{r}
mtcars$`Revolutions Per Minute`
```

Next, we calculate torque as follows:
```{r}
mtcars$`Torque` <- 
((mtcars$`Gross Horsepower`)*5252)/(mtcars$`Revolutions Per Minute`)
```

Again, we have another column consisting of quantitative data:
```{r}
mtcars$`Torque`
```

Next, we take two subsets of the newly updated dataset:
```{r}
cardata1 <- mtcars[c(15,4)]
cardata1
cardata2 <- mtcars[c(14,4)]
cardata2
```


We can also use data science to determine whether or not torque and gross horsepower are dependent of each other. Let us calculate the probability that a randomly selected model has a torque that is at least 5,000 and has a gross horsepower that is at least 200:
```{r}
#From this, the probability that a randomly selected model has a minimal torque of 5,000 is 22/32. We determine the answer by dividing the number of observation in this subset by the total number of observations in the original set.
c1 <- cardata1[cardata1$`Torque` >= 5000, c("Torque", "Gross Horsepower")]
str(c1)

#From this, the probability that a randomly selected model has a minimal gross horsepower of 200 is 7/32. We determine the answer the same way we determined it perivously.
c2 <- cardata1[cardata1$`Gross Horsepower` >= 200, c("Torque", "Gross Horsepower")]
str(c2)


#After multiplying the two probabilites we get an answer of 0.15. 


c3 <- cardata1[cardata1$`Torque` >= 5000 & cardata1$`Gross Horsepower` >= 200, c("Torque", "Gross Horsepower")]
str(c3)

#From this, the probability that a randomly selected model has both a minimal gross horsepower of 200 and a minimal torque of 5,000 is 7/32 or 0.22. Since this does not exactly equal the probability that we previously computed, we can conclude that torque and gross horsepower are dependent on each other. 

```

Let us do the same analysis that we did for the previous subset:

```{r}
#P(Revolutions Per Minute >= 110) = 14/32
c4 <- cardata2[cardata2$`Revolutions Per Minute` >= 110, c("Revolutions Per Minute", "Gross Horsepower")]
str(c4)

#P(Gross Horsepower >= 90) = 27/32
c5 <- cardata2[cardata2$`Gross Horsepower` >= 90, c("Revolutions Per Minute", "Gross Horsepower")]
str(c5)

#Multiplying the two probabilities we get 0.37.

#P(Revolutions Per Minute >= 110 AND Gross Horsepower >= 90) = 10/32 = 0.32
c6 <- cardata2[cardata2$`Gross Horsepower` >= 90 & cardata2$`Revolutions Per Minute` >= 110, c("Revolutions Per Minute", "Gross Horsepower")]
str(c6)

#In this case also, both quantities are dependent on each other. 
```

We want to find out if horsepower have a linear correlation with any of the calculated quantities. We answer the question by generating scatterplots, lines of best fit, and loess curves of best fit as follows:

```{r}
require(ggplot2)

ggplot(data = cardata1, 
aes(x = cardata1$`Torque`, 
y = cardata1$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth(method = lm)+
xlab("Torque")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Torque")

ggplot(data = cardata1, 
aes(x = cardata1$`Torque`, 
y = cardata1$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth()+
xlab("Torque")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Torque")

ggplot(data = cardata2, 
aes(x = cardata2$`Revolutions Per Minute`, 
y = cardata2$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth(method = lm)+
xlab("Revolutions Per Minute")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Revolutions Per Minute")

ggplot(data = cardata2, 
aes(x = cardata2$`Revolutions Per Minute`, 
y = cardata2$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth()+
xlab("Revolutions Per Minute")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Revolutions Per Minute")
```

Then, we calculate the correlation coefficients for both sets of data as follows:

```{r}
cor(cardata1$`Gross Horsepower`, cardata1$`Torque`)
cor(cardata2$`Gross Horsepower`, cardata2$`Revolutions Per Minute`)

#Horsepower has an almost-linear correlation with torque but not with the number of revolutions per minute. A line of best fit does exist in the scatter plot. The correlation coefficient with torque is 0.91 which is close to 1 which indicates that the correlation is almost linear. The correlation coefficient with the number of revolutions per minute is 0.06 which is close to 0 which indicates that there is almost no linear correlation with the number of revolutions per minute. It is almost impossible for a line of best fit to exist. However, it is possible for a loess smothed fit curve to exist for both relations.     
```

We generate frequency histograms for Gross Horsepower, Torque, and Revolutins Per Minute as follows:

```{r}
require(ggplot2)
qplot(cardata1$`Gross Horsepower`, data = cardata1, xlab = 'Gross Horsepower')
qplot(cardata1$`Torque`, data = cardata1, xlab = 'Torque')
qplot(cardata2$`Revolutions Per Minute`, data = cardata2, xlab = 'Revolutions Per Minute')
#None of the above distributions are normal. 
```

Here are the corresponding standard deviations for the above histograms:

```{r}
sd(cardata1$`Gross Horsepower`)
sd(cardata1$`Torque`)
sd(cardata2$`Revolutions Per Minute`)
#The data values for torque are the furthest spread from the mean because the standard deviation for that data set is equivalent to 3,383.843. 
```